---
title: "Midpoint"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("scripts/summary_info.R")
```

## Introduction Paragraph

For our project, we will be examining renewable energy information in Washington State with comparisons from data sets of national averages. In our summary information section, we include raw data regarding the electricity supply in Washington State and the energy consumption per state in the United States in 2018. With these data sets, we examine how committed the Washington Government is towards Renewable Energy, how Americans in general view it, and how Washingtonians see it in the summary section. We will turn this information into three data visualization including: Washington’s goal with the present energy levels, Washington’s renewable energy consumption compared to other states in the U.S., and the trend of renewable energy in Washington over the years.

## Summary
```{r get_summary_info, echo = FALSE}
info <- get_summary_info(elec_supply_wa, energy_consum_state)
```

### How committed is the Washington Government towards Renewable Energy?

Since 2001, Washington State had the highest share of renewable electricity
supply out of total electricity supply in
`r info$ren_elec_supply_wa_max$max_share_year` at
`r info$ren_elec_supply_wa_max$Share` percent, while only an year before,
it had the lowest share of renewable electricity supply out of total electricity
supply in `r info$ren_elec_supply_wa_min$min_share_year` at
`r info$ren_elec_supply_wa_min$Share` percent. This shows how quickly the share
of renewable electricity supply out of total electricity supply in Washington
State can change within a span of an year. Hence, to properly evaluate the
Washington State's commitment towards renewable energy, the average share over
2001-2018 gives a better indication than the highest and lowest amounts. At
`r info$avg_elec_ren_supply_share_wa$avg_share` percent, we can say that the
state's commitment towards renewable energy has been moderate, since the amount
is right in between the highest and lowest percentages.

### How do Americans view Renewable Energy?

In the United States, `r info$highest_ren_consum_us$highest_ren_state` had the
highest renewable energy consumption in 2018 at
`r info$highest_ren_consum_us$btus` BTUs,
while `r info$lowest_ren_consum_us$lowest_ren_state` had the lowest at
`r info$lowest_ren_consum_us$btus` BTUs. However, that's an unfair comparison
because densely populated states are likely to have a higher renewable energy
consumption than that of less densely populated states. Hence,
renewable energy consumption per total energy consumption would give us a better
understanding as to how the people of each state view renewable energy. Now,
`r info$high_ren_cons_per_total$highest_state` is the highest at
`r info$high_ren_cons_per_total$percent` percent, while
`r info$low_ren_cons_per_total$lowest_state` is still the lowest at
`r info$low_ren_cons_per_total$percent` percent.

### Where do Washingtonians stand in terms on Renewable Energy, relative to The United States?

The average US Renewable and Fossil Fuel Energy Consumption in 2018 was
`r info$consum_us$avgbtu_renew_consum` and
`r info$consum_us$avgbtu_fossil_consum` BTUs per state respectively. The
renewable energy consumption in Washington was
`r info$wa_consum$wa_total_ren_consum` BTUs, with the renewable
energy consumption per total energy consumption standing at
`r info$wa_consum$wa_ren_consum_per_total` percent. Here we can clearly see that
renewable energy consumption in Washington is much higher than the national
average, and renewable energy consumption per total energy consumption is only
slightly lower than that of Oregon, which has the highest. Hence, we can see 
that the Washington population is more than willing to move towards renewable
energy means and practices.
```{r, echo=FALSE, include=FALSE}
source("scripts/aggregate_table.R")
```

## Aggregated Data
#### Table 1: *Energy Consumption Percentages in WA 2000 - 2017*
```{r,echo=FALSE}
library("kableExtra")
kable(a, digits = 2, format = "html", row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                font_size = 12,
                position = "left") %>%
  scroll_box(height = "300px")
```
`*This table assumes that all forms of sun, wind, water, waste, and biomass energy sources are renewable.

Over the course of 17 years, Washington's renewable energy sources (excluding hydropower) generated <20% of the state's energy. The table reveals that without hydropower, Washington can only sustain <5% of Washington's energy requirement. This information is important as Gov. Jay Inslee pledged that Washington would achieve 100% carbon-neutral power by 2030. Without hydropower, Washington would have to look increase energy development efforts and look for alternative sources of power.



#### Table 2: *Energy Consumption Percentages in individual US States 2018*
This table assumes that all forms of sun, wind, water, and biomass energy sources are renewable.
```{r, echo=FALSE}
library("kableExtra")
kable(b, digits = 2, format = "html", row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                font_size = 12,
                position = "left") %>%
  scroll_box(height = "300px")
```

In 2018, the United States evaluated the amount of energy used by each state by British thermal units (BTU). There is no dominant renewable energy source type as energy produced by biomass and WWS are to some use in all states. The same idea applies to petroleum usage and other nonrenewable energy sources. The table reveals that nuclear energy is larger in larger states. In the entire continent, Mississippi (3.62%) had the least amount of its energy produced by renewable sources and Oregon (44.44%) with the most renewable energy.



#### Table 3: *Energy Consumption Percentages in the USA 2018*
```{r,echo=FALSE}
library("kableExtra")
kable(c, digits = 2, format = "html", row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped"),
                full_width = F,
                font_size = 12,
                position = "left") %>%
  scroll_box(height = "300px")
```
This table tracks the energy consumption per billion BTUs. Over the last 30 years it is observed that the US's renewable energy production fluctuates around the 20-30% range (see percentage renewable energy (out of total)). The renewable energy consupmtion grows relative to the increase in total energy consumption. Nuclear energy consumption is comparable to biomass energy consumption, whereas wind, water and sun energy (WWS) consumption is greater than both. The US may shift their focus to expanding multiple energy sources like nuclear energy and WWS to meet and hopefully completely fulfill the growing demand for energy.



## Three Charts

### 1. Scatter plot of Washington's Energy Usage per Type from 2001 - 2017
```{r, echo=FALSE}
source("scripts/chart_1.R")
suppressWarnings(plot(chart_1(energy_per_src_wa)))
```

The scatter plot precisely tracks the trend of both nonrenewable and renewable energy consumption in WA to see whether sufficients improvements have been made. We are making this chart to verify whether Washington is on track to achieve 100% carbon-neutral power by 2030. The plot depicts an increase in renewable energy sources, however, it flatlines from 2013-2017. Renewable energy points are seen to have a steady increase in 2015-2017. The trend line hints that renewable energy production is increasing, but steadily.

### 2. Grouped Bar plot of Washington vs US Average's renewable energy consumption
```{r, echo=FALSE}
source("scripts/chart_2.R")
chart2 <- chart_2(energy_consumption_per_state)
chart2
```

I used a grouped bar plot of Washington vs US average(without Washington)'s
renewable energy consumption since we wanted to know how Washington was doing
and if we were on course with our consumption of renewable energy compared to
the rest of the states.It could be easily observed that Washington on average
uses more renewable energy.

### 3. Map of renewable energy consumption percentage
```{r, echo=FALSE}
source("scripts/chart_3.R")
chart3 <- draw_map(energy_per_state)
chart3
```

I used a map to show the percentage renewable energy is being consumed by each
state since I thought it would be visually easier to see and compare each
state. It could be seen from the map that Washington has one of the highest
percentage of renewable energy usage in the United States. Each blue circle
has a pop-up which you can click to see the exact percentage.